Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2011-11-04 paper/self_calibration_I.tex
euclid: start adding rms, badness values from new results
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19862 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: adding results to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19863 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\subsection{Self-Calibration Performance}
The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets. With the Survey A dataset it does not converge, even when the system is started close to the optimum fit. The fitted instrument response solutions for the Survey D and A datasets are shown in Figures \ref{fig:D_result} and \ref{fig:A_result} respectively. The corresponding plot for the Survey B and C dataset has been omitted, due to its similarity to the Survey D dataset result. In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis''). The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.

@@

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{D_results.pdf}
\end{center}
\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models. (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response. (c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model. The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial). The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:D_result}}
\end{figure}

\clearpage
\begin{deluxetable}{ccccc}
\tablecaption{A summary of the quality of the final fit from the self-calibration procedure with the four simple survey strategies summarized in Table \ref{tab:simple_surveys}. The self-calibration procedure was run with only the brightest sources within the survey area; the source error $S_\text{RMS}$ corresponds to the measurements of these sources only and not all those within the survey. *Did not converge.\label{tab:simple_results}}
\tablehead{
\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $basis$ (\%)} }
\startdata
A*  & 8192 & XX & XX & XX \\
B  & 4160 & XX & XX & XX \\
C  & 26 & XX & XX & XX \\
D  & 9 & XX & XX & XX \\
\enddata
\end{deluxetable}

>>>>

\subsection{Self-Calibration Performance}
The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets. With the Survey A dataset it does not converge, even when the system is started close to the optimum fit. The fitted instrument response solutions for the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}. In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis''). The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.

@@

The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset. The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.

With the Survey C and D datasets, the instrument response is accurately recovered. As can be seen from Figs. \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$. The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the instrument response model fitted for.

The results from the self-calibration procedure with the Survey A and B datasets are also interesting.

tiling is crap

parts well connected work well

 when the fitted instrument response is compared to the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we can see that the final solution is close

possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{A_8192_ff.pdf}
\end{center}
\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models. The self-calibration procedure did not converged to the final solution, the results presented here correspond to the instrument response found after $8\,192$ iterations. (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response. (c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model. The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial). The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:A_result}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{B_4160_ff.pdf}
\end{center}
\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey B dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models. The self-calibration procedure converged to the final solution after 4160 iterations. (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response. (c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model. The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial). The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:B_result}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{C_026_ff.pdf}
\end{center}
\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey C dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models. The self-calibration procedure converged to the final solution after 26 iterations. (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response. (c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model. The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial). The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:C_result}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{D_009_ff.pdf}
\end{center}
\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models. The self-calibration procedure converged to the final solution after 9 iterations. (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response. (c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model. The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial). The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:D_result}}
\end{figure}

\clearpage
\begin{deluxetable}{ccccc}
\tablecaption{A summary of the quality of the final fit from the self-calibration procedure with the four simple survey strategies summarized in Table \ref{tab:simple_surveys}. The self-calibration procedure was run with only the brightest sources within the survey area; the source error $S_\text{RMS}$ corresponds to the measurements of these sources only and not all those within the survey. *Did not converge.\label{tab:simple_results}}
\tablehead{
\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $B_\basis$ (\%)} }
\startdata
A*  & 8192 & XX & XX & XX \\
B  & 4160 & XX & XX & XX \\
C  & 26 & 0.199 & 0.305 & 0.043 \\
D  & 9 & 0.113 & 0.304 & 0.032 \\
\enddata
\end{deluxetable}
